---
title: "Plots!"
author: "Maria Kamenetsky"
date: "Friday, October 30, 2015"
output: pdf_document
header-includes:
    \usepackage{graphicx}

---


#Graphs!


```{r eval=FALSE}
#1) Regular boxplots
q <- ggplot(df, aes(factor(Pal), SA))
q + geom_boxplot(outlier.colour="orange", outlier.size=3, aes(fill=factor(Pal))) +  coord_flip() + facet_grid(Dec ~ N)

#2) Boxplots with jittering
r <- ggplot(data=sub, aes(Conc,DevA))
r + geom_boxplot(outlier.colour="orange", outlier.size=6, aes(fill=Conc)) +
    geom_jitter() + coord_flip() + facet_grid(Pop~.) 

```


* _GeomPoint_
```{r eval=FALSE}
q <- ggplot(BA, aes(SA, DD))
q + geom_point() +  coord_flip() + facet_grid(Dec ~ Pal) #+  stat_smooth(method=lm, fullrange = TRUE) 

```


* _Histogram_
```{r eval=FALSE}
#1) histogram on log scale
ggplot(df, aes(LtoSM)) + geom_histogram() + scale_x_log10()

#2) Facetted Histogram
ggplot(svN, aes(sv, fill = reg)) + geom_histogram(binwidth=1) + facet_grid(rg ~ reg + soft) 


#3) Histogram
p <- qplot(mathkind, data=classroom, geom="histogram", 
           xlab="Mathematics Score in Kindergarten",main="Kindergarten Mathematics Score", 
           binwidth = 10)
p + geom_histogram(binwidth=10) 
```


* _qplot_
```{r eval=FALSE}
qplot(R1,it,  data=z)
```


* _statsmooth_
```{r eval=FALSE}
p <- ggplot(z, aes(R2, it))
p +  scale_y_continuous(limits=c(0,15)) +stat_smooth() 
#smoothing method (function) to use, eg. lm, glm, gam, loess, rlm. For datasets with n < 1000 default is loess. For datasets with 1000 or more observations defaults to gam, see gam for more details.
```



* _bar chart_
```{r eval=FALSE}
#Classic Barchart
qplot(factor(md), data=z, geom="bar", fill=factor(iG))

#Barchart with x-axis labels on 45 degree angle
q <-qplot(factor(md), data=z, geom="bar", fill=factor(reg)) 
q + theme(axis.text.x = element_text(angle = 45, hjust = 1))

#Facetted Bar Charts
ggplot(z, aes(md, fill=factor(reg))) + geom_bar() + facet_grid(rg ~ cf) + coord_flip()


```



* _Histograms_
```{r eval=FALSE}
#Facetted Histograms
ggplot(z, aes(md, fill=factor(reg))) + geom_bar() + facet_grid(rg ~ cf) + coord_flip()

```



* _Density Plots_
```{r eval=FALSE}
#1) Facetted Density Plots
ggplot(z, aes(salesdensity, fill = cf)) +  geom_density() +   scale_x_continuous(limits=c(0,4)) + facet_grid(cf ~ rg)

#2) Density Plots with 2 Factors - color
p <- qplot(mathkind, data=classroom, geom="density", 
           linetype=sex, color=sex,xlab="Mathematics Score in Kindergarten")
p + geom_density(aes(x=mathkind)) + labs(title="'Mathkind' by 'Sex'")
```



* _ScatterPlots_
```{r eval=FALSE}
#Scatter plot with point-wise confidence intervals
p <- qplot(mathgain, mathkind, data=classroom, geom=c("smooth"), 
           method="loess", xlab="Mathematics Gain for Students in Sample", 
           ylab="Mathematics Score in Kindergart.", 
           main="Scatterplot of 'Mathgain' vs. 'Mathkind'")
p + geom_point() + stat_smooth(se=TRUE) + 
  geom_smooth(method="loess") + 
  labs(title="'Mathgain' vs. 'Mathkind'")

#ScatterPlot with Regression Line (and no confidence bands)
p <- ggplot(classroom, aes(x=mathgain,y=mathkind), 
            geom_smooth(method="gam", formula = mathkind~mathgain), 
            stat_smooth(se=FALSE), 
            xlab="Mathematics Gain for Students in Sample",
            ylab="Mathematics Score in Kindergarten")
p + geom_point() + stat_smooth(se=FALSE) + 
  labs(title="Scatterplot of 'Mathgain' vs. 'Mathkind'") + 
  geom_smooth(method= "gam")

#Log Scatter Plot with GLM Smoother
v <- ggplot(ufc, aes(x=log(Height), y = log(Dbh))) + geom_point()
v + stat_smooth(se=TRUE) + geom_smooth(method="glm", formula = log(Dbh)~log(Height))

```



* _Dot Plot_

```{r eval=FALSE}
#Example 1
eleven <- subset(classroom, schoolid == "11")
p <- ggplot(eleven, aes(x=classid,y=mathgain),
            xlab="Classroom Identifier ", 
            ylab="Gain in Mathematics Score (from Kindergarten to First Grade")
p + geom_point() + labs(title="'Mathgain' by 'Classid' for SchoolID #11")


#Example 2
ggplot(ufc, aes(x=tr, y = Dbh)) + geom_point() + facet_wrap(~ sp)

```

# _GGMAP_
* _Getting the Map_
```{r eval=FALSE}
map <- get_map(location='united states', zoom=4, maptype='toner')
ggmap(map) + geom_point(aes(x=longitude, y= latitude, size=(it)), data=z, alpha=0.5)
#I supplied the long/lat here from my dataset
```

* _Plotting points_
```{r eval=FALSE}
#1) Basic Map
ggmap(map) + geom_point(aes(x=longitude, y= latitude, color=densityR1, fill=densityR1), data=z, alpha=0.5)

#2) Facetted Map (with colours specified)
ggmap(map) + geom_point(aes(x=longitude, y= latitude,  colour=it), size = 3, data=z, alpha=0.5) + facet_wrap(cf ~phR) + scale_colour_gradient(limits=c(3, 4), low="blue", high="red")

```



* _Keep Unit Aspect Ratio_
```{r eval=FALSE}
+ theme(aspect.ratio = 1)

```


* _Change Color Gradient and Point Size_
```{r eval=FALSE}
#1) Change color gradient Scale and specify which will be low
+ scale_colour_gradient(limits=c(3,4), low="red")

#2) Change with color brewer
+ scale_colour_brewer(type="div",palette="Set1")

#3) Point size (size=3)
ggmap(map) + geom_point(aes(x=longitude, y= latitude, color=sv), data=svN, colour="red", size=4)
```

* _density map_
```{r eval=FALSE}
z$cuts <- cut(z$it, breaks=5)
ggmap(map) + stat_density2d(aes(x=longitude, y= latitude, fill=cuts),
                            data=z,h=3 ,geom="polygon")
```




# _Maps with Spatial Polygons_
```{r eval=FALSE}
stP$name <- tolower(stP$stL)
stP <- stP[order(stP$name),] # reorder alpha
stP <- subset(stP, (name!="hawaii" & name!="alaska"))
us <- map("state", plot=FALSE, fill=TRUE)
us.ids <- sapply(strsplit(us$names,":"), function(x) x[1])
us.sp <- map2SpatialPolygons(us, us.ids, CRS("+proj=longlat + datum=wgs84"))

IDs <- match(stP$name, row.names(us.sp))
us.spNew <- us.sp[IDs]
row.names(stP) <- row.names(us.spNew) #these need to match
us.spdf <- SpatialPolygonsDataFrame(us.spNew, data=as.data.frame(stP)) # rownames of state need to be us.ids

#Plot
spplot(us.spdf, "reg", main="Title")

#Exclude Obs and Plot
exclude <- us.spdf[!(us.spdf$stsB==1888 | us.spdf$stsB==1734),]
sp5 <-spplot(exclude, "stsB",cuts=99, main="Title") 

```


* _Plotting Multiple ggplots into a Grid_
```{r eval=FALSE}
library(grid)
library(automap)
vplayout <- function(x, y) viewport(layout.pos.row = x, layout.pos.col = y)

plot1 <- ggplot(z, aes(phR,it)) + geom_boxplot(outlier.colour="orange", outlier.size=3, aes(fill=factor(phR))) +   scale_y_continuous(limits=c(0,75)) + coord_flip() + facet_grid(cf ~ soft) 
plot2 <- ggplot(z, aes(trR,it)) + geom_boxplot(outlier.colour="orange", outlier.size=3, aes(fill=factor(trR))) + scale_y_continuous(limits=c(0,75)) + coord_flip() + facet_grid(cf ~ soft) 
plot3 <- ggplot(z, aes(rrR,it)) + geom_boxplot(outlier.colour="orange", outlier.size=3, aes(fill=factor(rrR)))  + scale_y_continuous(limits=c(0,75)) + coord_flip() + facet_grid(cf ~ soft) 

grid.newpage()
pushViewport(viewport(layout = grid.layout(2,2)))
print(plot1, vp=vplayout(1,1))
print(plot2, vp=vplayout(1,2))
print(plot3, vp=vplayout(2,1))

```

##ggplot2 - General Tips and Tricks

* _Suppress Plotting NA's_
```{r eval=FALSE}
#outlier.shape=NA
r + geom_boxplot(outlier.shape=NA, aes(fill=factor(phR))) +  coord_flip() + scale_y_continuous(limits=c(0,2)) + coord_flip() + facet_grid(cf ~ soft) 
```

* _Setting Limits on X/Y Coords in ggplot_
```{r eval=FALSE}
#for y (or x) limits
+ scale_y_continuous(limits=c(0,50))

```

* _Dealing with NA's in DF to Plot_
```{r eval=FALSE}
#Solution - create sub data frame without NA's

#Example: use ggplot(na.omit(subset))
subz <- z[c("it","densityR1", "iG")]
b <- ggplot(na.omit(subz), aes(factor(densityR1), it))
b + geom_boxplot(aes(fill=factor(densityR1))) + scale_y_continuous(limits=c(0,100)) + coord_flip() + facet_grid(iG ~.)  #limit to 100


```



#Neat Random Tricks

* _Using texreg Package for Publication-Quality md Output Comparisons_
```{r eval=FALSE}
screenreg(list(mod11F, mod11IR, mod11P, mod11Ss)) #this makes output pretty on the screen
texreg(list(mod11F, mod11IR, mod11P, mod11Ss)) # this makes output pretty in latex
```


* _Other Tables_
```{r eval=FALSE}
xtabs(disease/total ~ sex+food, babyfood)

or xtable
or 
table

```


#Sweet R Packages

* {\bf tidyr} = some data wrangling things; has some SQL style things
* {\bf ggplot2} = literally the best
* {\bf gridExtra} = need this if you want to create panels with ggplot that aren't automated (where you can just facet-grid/wrap)
* {\bf ggmap} = mapping with ggmpa
* {\bf mapproj} = needed to get maps using spatial polygons
* {\bf stats} = used for function \textit{aggregate}
* {\bf texreg} = used for fancy tex output
* \bf{reshape} = this one makes more sense and is stata-like
* \bf{reshape2} = this one contains 'melt'
* \bf{xtable}







